A catalog of wide binaries and comoving star pairs from \textsl{Gaia DR1} and low-resolution spectroscopy

### 2017-07-03 paper/comoving-followup.tex
intro first pass done, woot

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, at least relative to the kinematic and chemical
dispersions observed amongst typical stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same initial chemistry
(e.g., cite + \citealt{Feng:2014}) and same ages.
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way; obvious dynamical
associations between stars from the same cluster are then erased as the
constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way.
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children so that they may be re-associated with precisely measured
abundance patterns (of $\gtrsim 10$ species).

The feasibility of chemical tagging depends on the intrinsic chemical
dispersions of stars formed from the same birth clusters, the precision of the
abundance measurements, and the ability to tell that the chemical abundances of
two stars are more similar than two randomly-paired Galactic disk stars.
Recent work suggests that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
In effect, for most elements, the intrinsic dispersions are unconstrained
because of measurement noise, so the true intrinsic dispersions are not known.
It was also recently shown that $\approx 0.3\%$ of random pairings of Galactic
field red giant stars have abundance patterns that are indistinguishable at the
same level (``doppelgangers''), even with 20 chemical species
(\citealt{Ness:2017}).

Are these properties of open clusters universal, or limited to the few clusters
presently studied?
Do the chemical properties of open clusters depend on the star formation rate or
efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses and member stars that span a large range in
masses and evolutionary stages.
\todo{Gaia will help \gaia\ (\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a})}
However, because they disrupt quickly, the present-day open cluster population
is biased towards younger or more massive clusters.

An alternate tracer of the chemical properties mentioned above are
widely-separated ($>XX~\pc$), coeval pairs or multiplets of stars.
These tracers include both bound and unbound wide-separation stellar systems; we
refer to these systems collectively as ``comoving stars'' because, at large
separations, with present-day and near-future velocity uncertainties, it is
typically not possible to determine whether these systems are in fact bound.
Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
dissolving or disrupted open clusters.
In any of these cases, it is likely that the stars formed from the same region
of a birth cloud, yet are separated enough that there has not been any
cross-pollution in the event of strong stellar winds or stellar evolution.

For chemical studies, open clusters have the advantage of having many stars per
cluster, but there are many fewer clusters than comoving stars.
Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
surveys (cite rave, etc.), several hundred such systems are already known
(\citealt{Shaya:2011}).
In addition, over ten-thousand \emph{candidate} comoving stars have been
discovered from astrometric data alone, using data from both the
\project{Hipparcos} mission (\citealt{hipparcos?}) and the \gaia\ mission
(\citealt{Oh:2017,Andrews:2017}).
- Need sample of confirmed widely-separated binaries, follow-up to get
  detailed abundances
- Need way to verify that widely separated stars are physically associated,
  via kinematics

In this work, we present a catalog of genuine comoving star pairs discovered
through radial velocity follow-up of candidate comoving stars discovered from
their astrometry alone (\citealt{Oh:2017}).
We expect high-resolution follow-up to study chemistry...

@@

\section{Results}

>>>>

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, at least relative to the kinematic and chemical
dispersions observed amongst typical stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same initial chemistry
(e.g., cite + \citealt{Feng:2014}) and same ages.
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way.
Obvious dynamical associations between stars from the same cluster are then
erased as the constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way (e.g., \citealt{todo}).
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children and preserved so that their initial association may be inferred
with precisely measured abundance patterns (of $\gtrsim 10$ species).

The feasibility of chemical tagging depends on the intrinsic chemical
dispersions of stars formed from the same birth clusters, the precision of the
abundance measurements, and the ability to tell that the chemical abundances of
two stars are more similar than two randomly-paired Galactic disk stars.
Recent work suggests that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
In effect, for most elements, the intrinsic dispersions are unconstrained
because of measurement noise, so the true intrinsic dispersions are not known.
This demonstration of chemical uniformity in open clusters is promising for
chemical tagging.
However, it was also recently shown that $\approx 0.3\%$ of random pairings of
Galactic field red giant stars have abundance patterns that are
indistinguishable at the same level (``doppelgangers''), even with 20 chemical
species (\citealt{Ness:2017}).

Are these properties of open clusters universal, or limited to the few clusters
presently studied?
Do the chemical properties of open clusters depend on the star formation rate or
efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses and member stars that span a large range in
masses and evolutionary stages.
Precise kinematic and color data from the \gaia\ mission
(\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
existing membership catalogs and will lead to discoveries of many more open
clusters and young star associations (see, e.g., \citealt{Oh:2017}).
However, because they disrupt quickly, the present-day open cluster population
is biased towards younger or more massive clusters.

An alternate tracer of the chemical properties mentioned above are
widely-separated ($>XX~\pc$), coeval pairs or multiplets of stars.
These tracers include both bound and unbound wide-separation stellar systems; we
refer to these systems collectively as ``comoving stars'' because, at large
separations with present-day and near-future velocity uncertainties, it is
typically not possible to determine whether these systems are truly bound.
Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
dissolving or disrupted open clusters.
In any of these cases, it is likely that the stars formed from the same region
of a birth cloud, yet are separated enough that there has not been any
cross-pollution in the event of strong stellar winds or stellar evolution.

For chemical studies, open clusters have the advantage of having many stars per
cluster, but there are many fewer clusters than comoving stars.
Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
surveys (cite rave, etc.), several hundred such systems are already known
(\citealt{Shaya:2011}).
In addition, over ten-thousand \emph{candidate} comoving stars have been
discovered from astrometric data alone, using data from both the
\project{Hipparcos} mission (\citealt{ESA:1997}---\citealt{Shaya:2011}) and
\tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
\citealt{Oh:2017,Andrews:2017}).
Radial velocity follow-up is necessary to confirm pairs using these astrometric
catalogs, though \gaia\ \DR{2} will provide radial velocities for the brightest
stars ($G < 12$).
High-resolution spectroscopy is still needed to measure detailed chemical
abundances of these pairs, which may present interesting surprises; one
large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
high-resolution, high signal-to-noise spectra display large abundance
differences ($\approx 0.2~\dex$) that appear to have resulted from massive
accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
submitted).

In this work, we present a catalog of genuine comoving star pairs discovered
through radial velocity follow-up of candidate comoving stars selected from
inferences using their astrometry alone (\citealt{Oh:2017}).
This work serves as a pilot study for data expected in upcoming data releases
from the \gaia\ mission.
We describe the data, observations, and spectroscopic reduction in
\sectionname~\ref{sec:data}.
If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
where we discuss our main results.

@@

\section{Results} \label{sec:results}
